Power simulation section

### 2023-08-31 docs/manuscript.Rmd
Update manuscript.Rmd
Changed intro and left comments

<!-- General intro, relevance of Human AI Interaction experiments, need for well-powered and analyzed experimental research in this field -->
With the growing integration of artificial intelligence (AI) into our daily lives and routines, it has become increasingly important to study and understand how humans interact with these intelligent systems. 
This is particularly relevant as AI becomes more sophisticated and is deployed to support decision-making processes in fields such as law and medicine that have a significant impact on individuals and society at large. 
Therefore, thorough experimental research is needed to study human-AI interaction to understand how people perceive and respond to these novel technologies. 
When conducting experimental research in this emerging field, it is essential to determine an appropriate sample size [@lakensSampleSizeJustification2022] to ensure that the results obtained are both robust and informative, and use appropriate statistical models to analyze the data.

<!-- Some basics on statistical power, power analysis and why it is important -->
In empirical research relying on hypothesis testing, the most common strategy to determine appropriate samples size is based on statistical power. 
Statistical power is defined as the probability that a hypothesis test of interest turns out significant, when analyzing repeated samples from a population with a true effect of some pre-specified size.
Less formally, power is also described as the probability that a hypothesis test accepts the alternative hypothesis $H_1$ if $H_1$ is indeed true.
Power analysis represents the act of calculating the statistical power for a given true effect and sample size. 
By running the calculation for a range of sample sizes, the required sample size to find an assumed true effect with desired statistical power can be determined.

<!-- Power simulations are hard! (especially for GLMMs) -->
Power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes needed to achieve sufficient statistical power for results with high evidencial value.
While user-friendly software is available for simple statistical models like t-tests, ANOVA, and linear regression, the increasing use of (generalized) linear mixed models (GLMMs; also called multilevel models) in human-AI interaction research introduces greater complexity to power analysis.
For GLMMs, analytic options to derive sample sizes are not available for general research designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022].
An increasing number of tutorials describe how to perform power simulation for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
However, most tutorials discuss only linear mixed models (LMMs) and focus on the most common designs, limiting the guidance available for researchers facing more complex study designs [@brownIntroductionLinearMixedEffects2021].
Moreover, existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that meta-analytic results or data from pilot studies are available to determine plausible values for all model parameters.
In contrast, simulations are often required, necessitating assumptions for all parameters in the model [@maxwellSampleSizePlanning2008].
Therefore, running sample size calculations is crucial for designing experiments that can provide meaningful insights into human-AI interaction.
Moreover, many journals and funding agencies now mandate the inclusion of power analysis as part of study protocols and grant proposals, recognizing their significance in ensuring robust and meaningful findings.
By incorporating GLMMs and power analysis into research planning, researchers can enhance the replicability and credibility of their work, ultimately contributing to the advancement of science [@yarkoniGeneralizabilityCrisis2022].

>>>>

<!-- General intro, relevance of Human AI Interaction experiments, need for well-powered and analyzed experimental research in this field -->
As artificial intelligence (AI) becomes more and more ubiquitous in our everyday lives and routines, it has become increasingly important to study and understand how humans interact with these intelligent systems. 
This is particularly relevant as AI models are getting more sophisticated and are deployed to support decision-making processes in areas such as law and medicine that have significant implications for individuals and society at large. 
Therefore, thorough research is needed to study the interaction between humans and AI technology to understand how people perceive, respond to and are influenced by these novel technologies. 
<!-- SG: I think that defining a problem is missing here. It could be that most computer scientists developing the models are not well-trained in experimental design, which results in many papers on human-AI interaction being poorly designed and executed. This is problematic as it waists resources and might even be dangerous when readers rely on results from "bad" research. -->
When planning experimental research in this emerging field, it is essential to determine an adequate sample size [@lakensSampleSizeJustification2022] to ensure that the results obtained are both robust and informative. Moreover, conducting a careful a priori power analysis helps researchers decide which experimental design and statistical models are both feasible and appropriate for answering their research questions and analysing the data. 
<!-- SG: I think the second point, "and use appropriate statistical models to analyze the data." needs further explanation. I tried to change it, but I am not sure if this is what you actually wanted to say-->
<!-- SG: Overall, I am not entirely sure if the last two sentences are super well placed here since the benefits of performing a power analysis will be repeated below. Please ignore this comment, when you think otherise-->

<!-- Some basics on statistical power, power analysis and why it is important -->
In empirical research relying on hypothesis testing, the most common strategy for determining an adequate sample size is based on statistical power. 
Statistical power is defined as the probability that a hypothesis test of interest <!-- is the "of interest" needed? --> turns out significant when analyzing repeated samples from a population with a true effect of some pre-specified size.
Less formally, power is also described as the probability that a hypothesis test accepts the alternative hypothesis $H_1$ if $H_1$ is indeed true.
A power analysis represents the act of calculating the statistical power for a given true effect and sample size. 
By running the calculation for a range of sample sizes, the required sample size can be determined so that researchers to find an assumed true effect with the desired statistical power.

<!-- Power simulations are hard! (especially for GLMMs) -->
Power analysis offers a valuable solution by allowing researchers to estimate the appropriate sample sizes required to achieve sufficient statistical power for results with high evidential value. <!-- SG: I am not 100% sure what the added benefit of this sentence is. It is pretty much what has been said in the previous sentence without the results with high evidential value". I would just add the last part to the previous sentence and then delete this one. -->
While user-friendly power analysis software is available for simple statistical models like t-tests, ANOVA, and linear regression, there are few tools are available for more sophisticated statistical methods. 
However, researchers studying human-AI interaction are increasingly using (generalized) linear mixed models (GLMMs; also called multilevel models), which add complexity to power analysis.
For GLMMs, analytic options to derive sample sizes are not available for general research designs [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022]. <!-- SG: I think I do not fully understand this sentence; do you mean interface-based software, or are you trying to focuse more on the flexibility of the research design/model specification? -->
A number of tutorials have been published describing how to perform power simulation for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021].
However, most of these tutorials only cover linear mixed models (LMMs) and focus on the most common designs. This narrow focus provides little guidance for researchers faced with more complex study designs, which are very common [@brownIntroductionLinearMixedEffects2021].
Moreover, existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters.
In contrast, simulations are often required, necessitating assumptions for all parameters in the model [@maxwellSampleSizePlanning2008].
Therefore, running sample size calculations is crucial for designing experiments that can provide meaningful insights into human-AI interaction.
Moreover, many journals and funding agencies now mandate the inclusion of power analysis as part of study protocols and grant proposals, recognizing their significance in ensuring robust and meaningful findings.
By incorporating GLMMs and power analysis into research planning, researchers can enhance the replicability and credibility of their work, ultimately contributing to the advancement of science [@yarkoniGeneralizabilityCrisis2022].
